Iterative Monomorphisation

### 2024-04-22 iterative_monomorphisation.tex
commented a table (preemptively)

% Provide complete list of problems explain which ones were excluded and why
% We ignore the following bounds: ty var limit per clause, subst per ty var, limiting the new type args per clause

@@

Each bound presented presents a tradeoff, a higher bound will allow for a more exhaustive monomorphisation at the cost of taking longer to run. Our main goal being higher performance automatic theorem provers, we chose the number of problems solved in a 30 second time-frame as a measure of success of the monomorphisation procedure. To find the best bounds for this, it would be ideal to exhaustively test all possible combinations of values each bound could take. Given the number of bounds, this is not realistic. Instead, closely linked bounds are grouped together and their possible values are more thouroughly searched. Once the best performing set of values for a group of bounds is found, these values are chosen as a default and the search for the next group begins.

\begin{table}[th]
\caption{Evaluation of bounds for monomorphic type argument generation}
\centering\begin{tabular}{@{}l*{9}{>{\centering\arraybackslash}p{1.1em}}@{}}
   \toprule
   & &&& \multicolumn{3}{c}{floor} \\
   & \multicolumn{3}{c}{1} & \multicolumn{3}{c}{3} & \multicolumn{3}{c}{9}\\
   \cmidrule(l){2-10}
   & &&& \multicolumn{3}{c}{multiplier} \\
   \multirow{1}{2em}{cap} & 0 & 1 & 2 & 0 & 1 & 2 & 0 & 1 & 2\\
    \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(l){8-10}
    25   & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    50   & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    100  & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \(\infty\) & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \bottomrule
\end{tabular}
\end{table}

@@

If the floor and multiplier are set to 0 and 0, no new polymorphic type arguments will be generated, as a result the cap will be irrelevant.

\begin{table}[th]
\caption{Evaluation of bounds for polymorphic type argument generation}
\centering\begin{tabular}{@{}l*{9}{>{\centering\arraybackslash}p{1.1em}}@{}}
   \toprule
   & &&& \multicolumn{3}{c}{floor} \\
   & \multicolumn{3}{c}{0} & \multicolumn{3}{c}{3} & \multicolumn{3}{c}{6}\\
   \cmidrule(l){2-10}
   & &&& \multicolumn{3}{c}{multiplier} \\
    \multirow{1}{2em}{cap} & 0 & 0.5 & 1 & 0 & 0.5 & 1 & 0 & 0.5 & 1 \\
    \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(l){8-10}
    10 &   & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    20 &   & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    40 &   & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \(\infty\) & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \bottomrule
\end{tabular}
\end{table}

@@

  * Böhme
  * similar algo implemented in Why3?

>>>>

To measure the performance of a Zipperposition instance, we chose as a measure of success the number of problems solved by that instance in under 30 seconds. This setup reflects practical use cases of automatic provers in the context of interactive provers. The problems used for these tests are taken from the TPTP library. The set of selected problems is made up of all problems in the TF1 and TH1 grammars, ie: all first-order and higher-order problems with rank-1 polymorphism. Because Zipperposition does not support reasoning with real numbers, all problems with real numbers or importing axioms with real numbers were removed.

@@

Each bound of the monomorphisation process presents a tradeoff: a higher bound will allow for a more exhaustive instantiations of type variables at the cost of taking more time. Ideally, all possible combinations of values for each bound could be exhaustively tested to find the best compromise between completeness and speed. Given the number of bounds, however, this is not feasible. Instead, closely related bounds are grouped together and combinations of values for bounds in these groups are tested. Once the best performing set of values for a group of bounds is found, these values are assigned to the corresponding bounds as the search for the next group begins. This method requires an initial default set of bounds, the values chosen for these are based on preliminary tests during development.

A potential pitfall of this approach is to overfit the bounds to the problems used as tests. To mitigate this, 500 test problems were taken at random from our set of 1754 selected TPTP problems. Tests whose results are used to inform the value of bounds or parameters always make use of the 500 problems set.

Table [ref] groups bounds that influence the maximum number of newly generated monomorphic type arguments per clause per iteration. It is calculated using three bounds [backwards ref] which therefore form a natural group. Our results show that in this configuration, there seems to be certain range within which the values of the bounds have little effect on performance. Beyond this range, however, the number of problems solved drops-off sharply.
\begin{table}[th]
\caption{Evaluation of bounds for monomorphic type argument generation}
\centering\begin{tabular}{@{}l*{9}{>{\centering\arraybackslash}p{1.5em}}@{}}
   \toprule
   & &&& \multicolumn{3}{c}{floor} \\
   & \multicolumn{3}{c}{50} & \multicolumn{3}{c}{100} & \multicolumn{3}{c}{200}\\
   \cmidrule(l){2-10}
   & &&& \multicolumn{3}{c}{multiplier} \\
   \multirow{1}{2em}{cap} & 0 & 1 & 2 & 0 & 1 & 2 & 0 & 1 & 2\\
    \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(l){8-10}
    250   & 164 & 164 & 164 & 164 & 164 & 163 & 163 & 164 & 164 \\
    500   & 164 & 163 & 164 & 164 & 165 & 163 & 164 & 164 & 164 \\
    750   & 164 & 164 & 165 & 164 & 163 & 164 & 164 & 164 & 164 \\
    %\(\infty\) & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \bottomrule
\end{tabular}
\end{table}

@@

Table [ref 2] fills the same role except it is for bounds limiting the number of new non-monomorphic type arguments. The bound values are lower because non-monomorphic type arguments tend to be more explosive. The results are similar to that of table 1 [ref ].

\begin{table}[th]
\caption{Evaluation of bounds for non-monomorphic type argument generation}
\centering\begin{tabular}{@{}l*{9}{>{\centering\arraybackslash}p{1.1em}}@{}}
   \toprule
   & &&& \multicolumn{3}{c}{floor} \\
   & \multicolumn{3}{c}{0} & \multicolumn{3}{c}{3} & \multicolumn{3}{c}{6}\\
   \cmidrule(l){2-10}
   & &&& \multicolumn{3}{c}{multiplier} \\
    \multirow{1}{2em}{cap} & 0 & 0.5 & 1 & 0 & 0.5 & 1 & 0 & 0.5 & 1 \\
    \cmidrule(lr){2-4} \cmidrule(lr){5-7} \cmidrule(l){8-10}
    10 &   & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    20 &   & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    40 &   & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \(\infty\) & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 & 0 \\
    \bottomrule
\end{tabular}
\end{table}

@@

  * Böhme
